---
title: "Finance Quantitative"
subtitle: "Calcul Actuariel"
author: Patrick Hénaff
date: "Version: `r format(Sys.Date(), '%d %b %Y')`"
output:
  pdf_document:
    keep_tex: false
    fig_caption: yes
    latex_engine: pdflatex
geometry: margin=1in

header-includes:
  - \usepackage[utf8]{inputenc}
  - \usepackage{eurosym}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load-libraries, include=FALSE, echo=TRUE}
library(quantmod)
library(xts)
library(hornpa)
library(lubridate)
library(xtable)
library(PerformanceAnalytics)
library(TTR)
library(SuppDists)
library(lubridate)
library(roll)
library(Hmisc)
library(nFactors)
library(kableExtra)
library(quadprog)
```

# Time Value of Money (Van Horne)

Consider the following cash flow streams:

```{=tex}
\begin{table}[ht]
  \centering
\begin{tabular}{llllll}
\toprule \\
Year & 1 & 2 & 3 & 4 & 5 \\
\midrule
W & 100 & 200 & 200 & 300 & 300 \\
X & 600 &     &     &     & \\
Y &     &     &     &     & 1200 \\
Z & 200 &     & 500 &     & 300 \\
\bottomrule
\end{tabular}
\end{table}
```
```{=tex}
\begin{enumerate}
  \item Compute the future value of each stream at a compounded rate of 10\%.
  \item Compute the present value of each stream at a compounded rate of 14\%.
\end{enumerate}
```
```{r}
CF = rbind(c(100, 200, 200, 300, 300), 
              c(600, 0, 0, 0, 0),
              c(0, 0, 0, 0, 1200), 
              c(200, 0, 500, 0, 300))
              
CF.fut = matrix(0, nrow = 4, ncol = 5)
CF.pst = matrix(0, nrow = 4, ncol = 5)

r.1 = 0.1
r.2 = 0.14

i = 0
for (i in 1:ncol(CF)){
  CF.fut[,i] <- CF[,i] * (1 + r.1)^(i)
  CF.pst[,i] <- CF[,i] / (1 + r.2)^(i)
}
  
print(CF.fut)
print(CF.pst)

```

# Comparing contracts (Van Horne)

On a contract, you have a choice of receiving 25,000 \euro in six years or 50,000 \euro in 12 years. At which compound annual rate would you be indifferent between the two options?

```{r}
amount.1 = 25000
amount.2 = 50000
y.1 = 6
y.2 = 12 

r.indiff = (amount.2/amount.1)^(1/(y.2 - y.1)) - 1
print(r.indiff)
```

# Mortgage (Van Horne)

You obtain a 10-year, 50,000 \euro  loan. The compound annual interest rate is 8%. The loan is paid back by 10 annual installments of 7,451.47 \euro.

```{=tex}
\begin{enumerate}
\item How much of the first year payment is principal?
\item How much total interest will be paid over the life of the loan?
\end{enumerate}
```
# Savings Plan

You need to have 50,000 \euro  at the end of ten years. To accumulate this sum, you plan to save a certain amount at the end of each year, for the next ten years. The bank pays 8% interest, compounded annually. How much should you save each year?

# Mortgage Paydown

You have borrowed 14,300 \euro  at a compound annual interest rate of 15%. You can make annual payments of 3,000 \euro  on your loan. How long will it be before your loan is completely paid down?
